---
title: "TBIJYemen"
author: "Soo Wan Kim"
date: "May 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(lubridate)
library(maptools)
library(raster)
library(rgdal)

theme_set(theme_minimal())

#saved xlsx file as csv in Excel
gtd <- read_csv("Data/Terrorism Data/globalterrorismdb_0616dist.csv")

gtd_filt <- gtd %>%
  mutate(TargetType = ifelse(targtype1_txt %in% c("Military", "Police", "Terrorists/Non-State Militia"),
"Military", 
ifelse(targtype1_txt %in% c("Government (Diplomatic)", "Government (General)", "Violent Political Party"),
       "Government", "Other"))) %>%
  mutate(WeaponType = ifelse(weaptype1_txt %in% c("Biological", "Chemical", "Radiological",
                                                  "Fake Weapons", "Sabotage Equipment",
                                                         "Vehicle (not to include vehicle-borne explosives, i.e., car or truck bombs)", "Other", "Unknown"), "Unknown/Other",
                                    ifelse(weaptype1_txt == "Explosives/Bombs/Dynamite", 
                                           "Explosives", weaptype1_txt)))
```

## Yemen

### Descriptive statistics

#### Terrorist attacks

Confirmed attacks
AQAP or Ansar al-Sharia

```{r}
gtd_yemen <- gtd_filt %>%
  filter(country_txt == "Yemen" & iyear > 2010 & guncertain1 == 0 & gname %in% 
           c("Al-Qaida in the Arabian Peninsula (AQAP)", "Al-Qaida in Yemen")) %>%
  mutate(date = paste0(iyear,"-",imonth,"-",iday)) %>%
  transform(date = as.Date(date, "%Y-%m-%d")) %>%
  arrange(date)
```

How many attacks?

When do attacks occur

```{r}
count_year <- gtd_yemen %>%
  group_by(iyear) %>%
  summarize(count = n())

ggplot(count_year, aes(iyear, count)) + 
  geom_line(linetype = 2) + 
  geom_point() + 
  labs(title = "Number of confirmed attacks by AQAP and affiliates, by year",
       x = "Year", y = "Number of attacks") + 
  theme(panel.grid.minor = element_blank())
```

Time between attacks (not counting attacks on the same day)

```{r}
gtd_yemen_casualties <- gtd_yemen %>%
  group_by(date) %>%
  summarize(attack_fatalities = sum(nkill))

date <- as.character(gtd_yemen_casualties$date)
date2 <- c(date, NA)
date2 <- date2[-1]
gtd_yemen_casualties$date2 <- as.Date(date2, "%Y-%m-%d")

gtd_yemen_casualties <- gtd_yemen_casualties %>%
  mutate(lull = date2 - date)

#histogram
ggplot(gtd_yemen_casualties, aes(log(as.numeric(lull) + 1))) + 
  geom_histogram(bins = 90)

#scatter plot over time
ggplot(gtd_yemen_casualties, aes(date, log(as.numeric(lull) + 1))) + 
  geom_point() + 
  geom_smooth()
```

Where do attacks occur

```{r}
yemen<-getData("GADM", country="YE", level=1)

yemen1 <-spTransform(yemen, CRS("+init=EPSG:2089"))

provinces <- unique(gtd_yemen$provstate)

gtd_yemen_map <- gtd_yemen %>%
  transform(provstate = ifelse(provstate == "Unknown", NA, provstate)) %>%
  transform(provstate = ifelse(provstate == "Adan", "`Adan", provstate)) %>%
  transform(provstate = ifelse(provstate == "Ad Dali", "Al Dali'", provstate)) %>%
  transform(provstate = ifelse(provstate == "Taizz", "Ta`izz", provstate)) %>%
  transform(provstate = ifelse(provstate == "Saada", "Sa`dah", provstate)) %>%
  transform(provstate = ifelse(provstate == "Sanaa", "San`a", provstate)) %>%
  transform(provstate = ifelse(provstate == "Marib", "Ma'rib", provstate)) 

#yemen1@data$NAME_1

NAME_1 <- yemen1@data$NAME_1

count_all <- gtd_yemen_map %>%
  group_by(provstate) %>%
  summarize(count = n()) %>%
  rename(`NAME_1` = provstate)

yemen1@data$id <- rownames(yemen1@data)
yemen1@data <- left_join(yemen1@data, count_all, by = "NAME_1")
yemen1_df <- fortify(yemen1)
yemen1_df <- left_join(yemen1_df, yemen1@data, by = "id")

theme_opts<-list(theme(panel.grid.minor = element_blank(),
                       panel.grid.major = element_blank(),
                       panel.background = element_blank(),
                       plot.background = element_blank(),
                       axis.line = element_blank(),
                       axis.text.x = element_blank(),
                       axis.text.y = element_blank(),
                       axis.ticks = element_blank(),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank()))

ggplot() + 
  geom_polygon(data = yemen1_df, aes(x = long, y = lat, group = group, fill =
                                       count), color = "black", size = 0.25) +
  theme(aspect.ratio=.7) + 
  labs(title = "Confirmed terrorist attacks by AQAP and affiliates, 2011 - 2015",
       fill = "Frequency") + 
  scale_fill_gradient(low='#FFFFCC', high='#006837') + 
  theme_opts
```

How many killed in attacks
`r sum(gtd_yemen$nkill, na.rm = TRUE)`
This field stores the number of total confirmed fatalities for the incident. The
number includes all victims and attackers who died as a direct result of the incident.

```{r}
fatalities <- gtd_yemen %>%
  group_by(iyear) %>%
  summarize(NumAttacks = n(),
            Sum = sum(nkill, na.rm = TRUE), 
            Min = min(nkill, na.rm = TRUE), 
            Max = max(nkill, na.rm = TRUE), 
            Mean = mean(round(nkill,2), na.rm = TRUE),
            Median = median(nkill, na.rm = TRUE),
            Sd = sd(round(nkill,2), na.rm = TRUE))

ggplot(gtd_yemen, aes(nkill)) + 
  geom_histogram(bins = 80)

ggplot(gtd_yemen, aes(date, log(nkill + 1))) + 
  geom_point() + 
  geom_smooth(color = "#006400")
```

Who killed

```{r}
targ_count <- gtd_yemen %>%
  group_by(iyear, TargetType) %>%
  summarize(freq = n()) %>%
  mutate(freq_weight = freq/sum(freq)) %>%
  transform(TargetType = as.factor(TargetType)) %>%
  transform(TargetType = factor(targ_count$TargetType, levels = c("Other", "Government", "Military")))

ggplot(targ_count, aes(iyear, freq_weight)) + 
  geom_bar(aes(fill = TargetType), stat = "identity") + 
  labs(title = "Percentage of attacks by target type",
       x = "Year", y = "Percentage of attacks",
       fill = "Target Type") + 
  scale_y_continuous(labels = percent_format())
```

how killed

```{r}
weap_count <- gtd_yemen %>%
  group_by(iyear, WeaponType) %>%
  summarize(freq = n()) %>%
  mutate(freq_weight = freq/sum(freq)) %>%
  transform(WeaponType = as.factor(WeaponType)) %>%
  transform(WeaponType = factor(weap_count$WeaponType, 
                                levels = c("Unknown/Other", "Incendiary", "Melee",
                                           "Firearms", "Explosives")))

ggplot(weap_count, aes(iyear, freq_weight)) + 
  geom_bar(aes(fill = WeaponType), stat = "identity") + 
  labs(title = "Percentage of attacks by weapon type",
       x = "Year", y = "Percentage of attacks",
       fill = "Target Type") + 
  scale_y_continuous(labels = percent_format())

sui_count <- gtd_yemen %>%
  group_by(iyear, suicide) %>%
  summarize(freq = n()) %>%
  mutate(freq_weight = freq/sum(freq))

ggplot(sui_count, aes(iyear, freq_weight)) + 
  geom_bar(aes(fill = suicide), stat = "identity")
  
```

suicide

```{r}

```


#### Covert strikes

```{r}
TBIJYemen <- read_csv("Data/Strikes Data/Tidy data/TBIJYemenTidy.csv") %>%
  transform(Date = as.Date(Date, "%Y-%m-%d")) %>%
  transform(Province = ifelse(Province %in% c("Hadhramout", "Hadramout", "Hadaramout"), 
                              "Hadramawt", 
                              ifelse(Province %in% c("Across Yemen", "Southern Yemen", "Between Marib and Shabwa", "Shabwa/Hadramout", "Shabwa/Mareb border", "Jawf-Mareb border", "Bayda-Shabwa border", "Central Yemen", "-", "Marib/Al Jawf", "Across central Yemen", "Yemen", "Multiple provinces"), NA, Province)))
TBIJYemen$Province <- gsub("Sanaa", "San`a", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Shabwa", "Shabwah", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Saada", "Sa`dah", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Marib", "Ma'rib", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Mareb", "Ma'rib", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Jawf", "Al Jawf", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Bayda", "Al Bayda", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Damar", "Dhamar", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Lahj", "Lahij", TBIJYemen$Province)

NAFYemen <- read_csv("Data/Strikes Data/Tidy data/NAFYemenTidy.csv")
NAFYemen <- NAFYemen[c("Date", "NumLeadersKilled")]

TBIJYemen <- left_join(TBIJYemen, NAFYemen, by = "Date") %>%
  mutate(LeaderKilled = ifelse(!is.na(NumLeadersKilled) & NumLeadersKilled > 0, 1, 0))
```

strike frequency by year

```{r}
strike_count_year <- TBIJYemen %>%
  mutate(year = year(Date)) %>%
  group_by(year) %>%
  summarize(count = n())

ggplot(strike_count_year, aes(year, count)) + 
  geom_line(linetype = 2) + 
  geom_point() + 
  scale_x_continuous(breaks = c(2002, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017)) + 
  labs(title = "Frequency of covert strikes in Yemen, by year",
       x = "Year", y = "Number of strikes",
       caption = "The Bureau of Investigative Journalism") + 
  theme(panel.grid.minor = element_blank())
```


location of strikes

```{r}
yemen2 <- yemen

count_strikes <- TBIJYemen %>%
  group_by(Province) %>%
  summarize(count = n()) %>%
  rename(`NAME_1` = Province)

yemen2@data$id <- rownames(yemen2@data)
yemen2@data <- left_join(yemen2@data, count_strikes, by = "NAME_1")
yemen2_df <- fortify(yemen2)
yemen2_df <- left_join(yemen2_df, yemen2@data, by = "id")

ggplot() + 
  geom_polygon(data = yemen2_df, aes(x = long, y = lat, group = group, fill =
                                       count), color = "black", size = 0.25) +
  theme(aspect.ratio=.7) + 
  labs(title = "Location of covert strikes on AQAP and affiliates in Yemen, 2002 - 2017",
       fill = "Frequency", caption = "The Bureau of Investigative Journalism") + 
  scale_fill_gradient(low='#FFFFCC', high='#006837') + 
  theme_opts
```

number killed in strikes

```{r}
fatalities <- gtd_yemen %>%
  group_by(iyear) %>%
  summarize(NumAttacks = n(),
            Sum = sum(nkill, na.rm = TRUE), 
            Min = min(nkill, na.rm = TRUE), 
            Max = max(nkill, na.rm = TRUE), 
            Mean = mean(round(nkill,2), na.rm = TRUE),
            Median = median(nkill, na.rm = TRUE),
            Sd = sd(round(nkill,2), na.rm = TRUE))

ggplot(gtd_yemen, aes(nkill)) + 
  geom_histogram(bins = 80)

ggplot(gtd_yemen, aes(date, log(nkill + 1))) + 
  geom_point() + 
  geom_smooth(color = "#006400")
```


leaders killed

```{r}
year <- sort(rep(c(2002, seq(2009, 2017, 1)), 2))
LeaderKilled <- rep(c(0,1), 10)
decap <- data.frame(year, LeaderKilled)

decap_count <- TBIJYemen %>%
  group_by(year(Date), LeaderKilled) %>%
  summarize(freq = n()) %>%
  rename(year = `year(Date)`)

decap_count <- left_join(decap, decap_count, by = c("year", "LeaderKilled")) %>%
  transform(freq = ifelse(is.na(freq), 0, freq))

decap_count <- left_join(decap_count, strike_count_year, by = "year") %>%
  mutate(freq_weight = freq/count) %>%
  transform(LeaderKilled = as.factor(LeaderKilled))

ggplot(decap_count, aes(year, freq_weight)) + 
  geom_bar(aes(fill = LeaderKilled), stat = "identity") + 
  labs(title = "Percentage of strikes known to have killed at least one group leader",
       x = "Year", y = "Percentage of attacks",
       fill = "Killed leader?",
       caption = "Sources: The Bureau of Investigative Journalism,\nNew America Foundation") + 
  scale_y_continuous(labels = percent_format()) + 
  scale_x_continuous(breaks = c(2002, 2009, 2011, 2013, 2015, 2017)) + 
  scale_fill_manual(labels = c("No", "Yes"), values = c("black", "#999999"))
```



### Regression analysis

```{r}
#attacks on same day treated as single attack

d <- 4/5
a <- 0.2
r <- 0.9


#more variables: current month, prev month, current and previous month smoothed

get_strike_casualties <- function(date) {
  obs <- filter(TBIJYemen, Date < date)
  obs <- group_by(obs, Date)
  obs <- summarize(obs, TotalKilledMin = sum(TotalKilledMin), LeaderKilled = sum(LeaderKilled))
  obs <- mutate(obs, Diff = ceiling((date - Date)/30))
  obs <- mutate(obs, TotKilledMinW = ifelse(LeaderKilled == 0,
                                            r*TotalKilledMin*d^(as.numeric(Diff) - 1),
                                            r*(1 + a)*TotalKilledMin*d^(as.numeric(Diff) - 1)))
  TotKilledMin <- sum(obs$TotKilledMinW)
  return(TotKilledMin)
}

TotKilledMinW <- map(gtd_yemen_casualties$date, function(i) get_strike_casualties(i))
TotKilledMinW <- data.frame(TotKilledMinW)
TotKilledMinW <- t(TotKilledMinW)
TotKilledMinW <- data.frame(TotKilledMinW)
rownames(TotKilledMinW) <- NULL

gtd_yemen_casualties <- bind_cols(gtd_yemen_casualties, TotKilledMinW)

ggplot(gtd_yemen_casualties, aes(TotKilledMinW, lull)) + 
  geom_point() + 
  geom_smooth()

# number of strikes in same month
# number of strikes in previous month
# smoothed (exponential?)
```

```{r}  
x <- seq(1, 1000, 1)
y <- 0

for (i in 1:1000) {
  y[i] <- d^(i - 1)
}

test <- data.frame(x, y)

ggplot(test, aes(x, y)) + 
  geom_point() + 
  xlim(0,50)

```

## Somalia

Intro

### Descriptive statistics

#### Terrorist attacks

#### Covert strikes

### Regression analysis