---
title: 'Extension: Fatalities'
author: "Soo Wan Kim"
date: "May 27, 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(lubridate)
library(maptools)
library(raster)
library(rgdal)
library(broom)
library(modelr)
library(knitr)

theme_set(theme_minimal())

#saved xlsx file as csv in Excel
gtd <- read_csv("Data/Terrorism Data/globalterrorismdb_0616dist.csv")

gtd_filt <- gtd %>%
  mutate(TargetType = ifelse(targtype1_txt %in% c("Military", "Police", "Terrorists/Non-State Militia"),
"Military", 
ifelse(targtype1_txt %in% c("Government (Diplomatic)", "Government (General)", "Violent Political Party"),
       "Government", "Other"))) %>%
  mutate(WeaponType = ifelse(weaptype1_txt %in% c("Biological", "Chemical", "Radiological",
                                                  "Fake Weapons", "Sabotage Equipment",
                                                         "Vehicle (not to include vehicle-borne explosives, i.e., car or truck bombs)", "Other", "Unknown"), "Unknown/Other",
                                    ifelse(weaptype1_txt == "Explosives/Bombs/Dynamite", 
                                           "Explosives", weaptype1_txt)))
```

## Yemen

### Descriptive statistics

#### Terrorist attacks

I look at all terrorist attacks in Yemen between the beginning of 2011 and the end of 2015 where the perpetrating group is coded in GTD as AQAP or Al-Qaeda in Yemen. In addition, I look only at observations where the attribution of the attack to the group is not based on speculation or dubious scoures (`guncertain1` = 0 in GTD).

```{r}
gtd_yemen <- gtd_filt %>%
  filter(country_txt == "Yemen" & iyear >= 2011 & guncertain1 == 0 & gname %in% 
           c("Al-Qaida in the Arabian Peninsula (AQAP)", "Al-Qaida in Yemen")) %>%
  mutate(date = paste0(iyear,"-",imonth,"-",iday)) %>%
  transform(date = as.Date(date, "%Y-%m-%d")) %>%
  arrange(date)
```

How many attacks?

When do attacks occur

```{r}
count_year <- gtd_yemen %>%
  group_by(iyear) %>%
  summarize(count = n())

ggplot(count_year, aes(iyear, count)) + 
  geom_line(linetype = 2) + 
  geom_point() + 
  labs(title = "Number of terrorist incidents carried out by AQAP and affiliates, by year",
       x = "Year", y = "Number of attacks") + 
  theme(panel.grid.minor = element_blank())
```

Time between attacks (not counting attacks on the same day)

```{r}
gtd_yemen_casualties <- gtd_yemen %>%
  group_by(date) %>%
  summarize(attack_fatalities = sum(nkill))

date <- as.character(gtd_yemen_casualties$date)
date2 <- c(date, NA)
date2 <- date2[-1]
gtd_yemen_casualties$date2 <- as.Date(date2, "%Y-%m-%d")

gtd_yemen_casualties <- gtd_yemen_casualties %>%
  mutate(lull = date2 - date)

#scatter plot over time
ggplot(gtd_yemen_casualties, aes(date, log(as.numeric(lull) + 1))) + 
  geom_point(alpha = 0.3) + 
  geom_smooth() + 
  labs(title = "Intervals between attacks, logged",
       x = "Date", y = "Intervals (logged)")
```

Who killed

```{r}
targ_count <- gtd_yemen %>%
  group_by(iyear, TargetType) %>%
  summarize(freq = n()) %>%
  mutate(freq_weight = freq/sum(freq)) %>%
  transform(TargetType = as.factor(TargetType)) %>%
  transform(TargetType = factor(targ_count$TargetType, levels = c("Other", "Government", "Military")))

ggplot(targ_count, aes(iyear, freq_weight)) + 
  geom_bar(aes(fill = TargetType), stat = "identity") + 
  labs(title = "Percentage of attacks by target type",
       x = "Year", y = "Percentage of attacks",
       fill = "Target Type") + 
  scale_y_continuous(labels = percent_format())
```

how killed

```{r}
weap_count <- gtd_yemen %>%
  group_by(iyear, WeaponType) %>%
  summarize(freq = n()) %>%
  mutate(freq_weight = freq/sum(freq)) %>%
  transform(WeaponType = as.factor(WeaponType)) %>%
  transform(WeaponType = factor(weap_count$WeaponType, 
                                levels = c("Unknown/Other", "Incendiary", "Melee",
                                           "Firearms", "Explosives")))

ggplot(weap_count, aes(iyear, freq_weight)) + 
  geom_bar(aes(fill = WeaponType), stat = "identity") + 
  labs(title = "Percentage of attacks by weapon type",
       x = "Year", y = "Percentage of attacks",
       fill = "Target Type") + 
  scale_y_continuous(labels = percent_format())

sui_count <- gtd_yemen %>%
  group_by(iyear, suicide) %>%
  summarize(freq = n()) %>%
  mutate(freq_weight = freq/sum(freq))

ggplot(sui_count, aes(iyear, freq_weight)) + 
  geom_bar(aes(fill = suicide), stat = "identity")
  
```

suicide

```{r}

```


#### Covert strikes

```{r}
TBIJYemen <- read_csv("Data/Strikes Data/Tidy data/TBIJYemenTidy.csv") %>%
  transform(Date = as.Date(Date, "%Y-%m-%d")) %>%
  transform(Province = ifelse(Province %in% c("Hadhramout", "Hadramout", "Hadaramout"), 
                              "Hadramawt", 
                              ifelse(Province %in% c("Across Yemen", "Southern Yemen", "Between Marib and Shabwa", "Shabwa/Hadramout", "Shabwa/Mareb border", "Jawf-Mareb border", "Bayda-Shabwa border", "Central Yemen", "-", "Marib/Al Jawf", "Across central Yemen", "Yemen", "Multiple provinces"), NA, Province)))
TBIJYemen$Province <- gsub("Sanaa", "San`a", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Shabwa", "Shabwah", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Saada", "Sa`dah", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Marib", "Ma'rib", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Mareb", "Ma'rib", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Jawf", "Al Jawf", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Bayda", "Al Bayda", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Damar", "Dhamar", TBIJYemen$Province)
TBIJYemen$Province <- gsub("Lahj", "Lahij", TBIJYemen$Province)

TBIJYemen <- TBIJYemen %>%
  mutate(TotalKilledMed = floor((TotalKilledMin + TotalKilledMax)/2)) %>%
  mutate(CiviliansKilledMed = floor((CiviliansKilledMin + CiviliansKilledMax)/2)) %>%
  mutate(MilitantsKilledMed = TotalKilledMed - CiviliansKilledMed)

#Leaders killed
NAFYemenLeaders <- read_csv("Data/Strikes Data/Tidy data/NAFYemenTidy.csv") %>%
  group_by(Date) %>%
  summarize(NumLeadersKilled = sum(NumLeadersKilled)) %>%
  transform(NumLeadersKilled = ifelse(Date == "2012-05-10", 2, 
                                      ifelse(Date == "2012-05-28", 1, 
                                             ifelse(Date == "2012-12-24", 1,
                                                    ifelse(Date == "2017-03-02", 1,
                                                           NumLeadersKilled))))) %>%
  mutate(LeaderKilled = ifelse(NumLeadersKilled > 0, 1, 0))
```

strike frequency by year

```{r}
strike_count_year <- TBIJYemen %>%
  mutate(year = year(Date)) %>%
  group_by(year) %>%
  summarize(count = n())

ggplot(strike_count_year, aes(year, count)) + 
  geom_line(linetype = 2) + 
  geom_point() + 
  scale_x_continuous(breaks = c(2002, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017)) + 
  labs(title = "Frequency of covert strikes in Yemen, by year",
       x = "Year", y = "Number of strikes",
       caption = "The Bureau of Investigative Journalism") + 
  theme(panel.grid.minor = element_blank())
```


number killed in strikes

define "incident"

```{r}
fatalities <- gtd_yemen %>%
  group_by(iyear) %>%
  summarize(NumAttacks = n(),
            Sum = sum(nkill, na.rm = TRUE), 
            Min = min(nkill, na.rm = TRUE), 
            Max = max(nkill, na.rm = TRUE), 
            Mean = mean(round(nkill,2), na.rm = TRUE),
            Median = median(nkill, na.rm = TRUE),
            Sd = sd(round(nkill,2), na.rm = TRUE))

strike_fatalities <- TBIJYemen %>%
  filter(year(Date) > 2002) %>%
  group_by(Date) %>%
  summarize(TotalMed = sum(TotalKilledMed),
            CiviliansMed = sum(CiviliansKilledMed),
            MilitantsMed = sum(MilitantsKilledMed)) %>% 
  gather(TotalMed, CiviliansMed, MilitantsMed, key = "type", value = "killed")
  

ggplot(strike_fatalities, aes(killed)) + 
  geom_density(aes(color = type))

ggplot(strike_fatalities, aes(log(killed + 1))) + 
  geom_density(aes(color = type))

ggplot(strike_fatalities, aes(Date, killed)) + 
  geom_point(aes(color = type)) + 
  geom_smooth(aes(color = type))
```


leaders killed

```{r}
year <- sort(rep(c(2002, seq(2009, 2017, 1)), 2))
LeaderKilled <- rep(c(0,1), 10)
decap <- data.frame(year, LeaderKilled)

decap_count <- NAFYemenLeaders %>%
  group_by(year(as.Date(Date, "%Y-%m-%d")), LeaderKilled) %>%
  summarize(freq = n())
colnames(decap_count) <- c("year", "LeaderKilled", "freq")

decap_count <- left_join(decap, decap_count, by = c("year", "LeaderKilled")) %>%
  transform(freq = ifelse(is.na(freq), 0, freq))

decap_count <- left_join(decap_count, strike_count_year, by = "year") %>%
  mutate(freq_weight = freq/count) %>%
  transform(LeaderKilled = as.factor(LeaderKilled))

ggplot(decap_count, aes(year, freq_weight)) + 
  geom_bar(aes(fill = LeaderKilled), stat = "identity") + 
  labs(title = "Percentage of strikes known to have killed at least one group leader",
       x = "Year", y = "Percentage of attacks",
       fill = "Killed leader?",
       caption = "Sources: The Bureau of Investigative Journalism,\nNew America Foundation") + 
  scale_y_continuous(labels = percent_format()) + 
  scale_x_continuous(breaks = c(2002, 2009, 2011, 2013, 2015, 2017)) + 
  scale_fill_manual(labels = c("No", "Yes"), values = c("black", "#999999"))
```



### Regression analysis

```{r}
#attacks on same day treated as single attack

d <- 4/5
a <- 0.1
r <- 0.05

#more variables: current month, prev month, current and previous month smoothed

get_strike_casualties <- function(date) {
  obs <- filter(TBIJYemen, Date < date)
  obs <- group_by(obs, Date)
  obs <- summarize(obs, TotalKilledMed = sum(TotalKilledMed))
  leaders <- filter(NAFYemenLeaders, as.Date(Date, "%Y-%m-%d") < date)
  obs <- left_join(obs, leaders, by = "Date")
  obs <- transform(obs, NumLeadersKilled = ifelse(is.na(NumLeadersKilled), 0, NumLeadersKilled))
  obs <- mutate(obs, Diff = ceiling((date - Date)/30))
  obs <- mutate(obs, TotKilledMedW = (1-r)*((1 + a)^(NumLeadersKilled))*TotalKilledMed*d^(as.numeric(Diff) - 1))
  TotKilledMed <- sum(obs$TotKilledMedW)
  return(TotKilledMed)
}

TotKilledMedW <- map(gtd_yemen_casualties$date, function(i) get_strike_casualties(i))
TotKilledMedW <- data.frame(TotKilledMedW)
TotKilledMedW <- t(TotKilledMedW)
TotKilledMedW <- data.frame(TotKilledMedW)
rownames(TotKilledMedW) <- NULL

gtd_yemen_casualties2 <- bind_cols(gtd_yemen_casualties, TotKilledMedW)

ggplot(gtd_yemen_casualties2, aes(log(TotKilledMedW + 1), log(as.numeric(lull) + 1))) + 
  geom_point() + 
  geom_smooth()

ggplot(gtd_yemen_casualties2, aes(log(TotKilledMedW + 1), log(attack_fatalities + 1))) + 
  geom_point() + 
  geom_smooth()

ggplot(gtd_)
```

```{r}
yemen_lm_data <- gtd_yemen_casualties2 %>%
  na.omit()

time_mod <- lm(log(as.numeric(lull)+1) ~ log(TotKilledMedW+1) + (log(TotKilledMedW+1))^2, data = yemen_lm_data)

tidy(time_mod) %>%
  kable(format = "html")

killed_mod <- lm(log(attack_fatalities+1) ~ log(TotKilledMedW+1) + (log(TotKilledMedW+1))^2, data = yemen_lm_data)

tidy(killed_mod) %>%
  kable(format = "html")

gtd_yemen_casualties2 %>%
  add_predictions(time_mod) %>%
  add_residuals(time_mod) %>%
  ggplot(aes(pred, resid)) +
  geom_point(alpha = .2) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_quantile(method = "rqss", lambda = 5, quantiles = c(.05, .95)) +
  labs(title = "Variance of error terms",
       x = "Predicted values",
       y = "Residuals")

time_augment <- yemen_lm_data %>%
  mutate(hat = hatvalues(time_mod),
         student = rstudent(time_mod),
         cooksd = cooks.distance(time_mod))

gtd_time_inf <- time_augment %>%
  filter(cooksd > 4 / (nrow(.) - (length(coef(time_mod)) - 1) - 1))

gtd_time_n_inf <- time_augment %>%
  filter(cooksd < 4 / (nrow(.) - (length(coef(time_mod)) - 1) - 1))
gtd_time_n_inf <- gtd_time_n_inf[,1:5]

time_mod2 <- lm(log(as.numeric(lull)+1) ~ log(TotKilledMedW+1) + (log(TotKilledMedW+1))^2, data = gtd_time_n_inf)

tidy(time_mod2) %>%
  kable(format = "html")

killed_augment <- yemen_lm_data %>%
  mutate(hat = hatvalues(killed_mod),
         student = rstudent(killed_mod),
         cooksd = cooks.distance(killed_mod))

gtd_killed_n_inf <- killed_augment %>% 
  filter(cooksd < 4 / (nrow(.) - (length(coef(killed_mod)) - 1) - 1))
gtd_killed_n_inf <- gtd_killed_n_inf[,1:5]
  
killed_mod2 <- lm(log(attack_fatalities+1) ~ log(TotKilledMedW+1) + (log(TotKilledMedW+1))^2, data = gtd_killed_n_inf)

tidy(killed_mod2) %>%
  kable(format = "html")
```

